---
title: "lab_6"
author: "Gabriel De La Rosa"
date: "2/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Attach Packages

```{r}
library(tidyverse)
library(here)
library(janitor)
library(raster)
library(sf)
library(tmap)
library(tmaptools)
library(gstat)
```

## Look at Grand Canyon GeoTIFF

```{r}
gc_dem <- raster(here("data", "gc_dem.tif"))

# Look at this using plot()

plot(gc_dem)

# Let's check the CRS @ = $ 

gc_dem@crs

# UTM projection is in meters (this kinda sucks)
# Check the extent (bounding box)

gc_dem@extent

# Lets change the projection to lat/long
# Copy the CRS arguments after running gc_dem@crs

# Create a wgs84 with latlong:

wgs_84 <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs"

gc_reproj <- projectRaster(gc_dem, crs = wgs_84, method = "bilinear")

gc_reproj@extent
```

 ### Crop raster to a smaller area (bounding box) ******
 
```{r}
bounds <- as(extent(-112.4, -112.0, 36.1, 36.3), 'SpatialPolygons')

```
 
### Make the CRS of our bounding box the same as for gc_reproj

```{r}
crs(bounds) <- crs(gc_reproj)
```

### Now, let's crop our original spatial data to the bounding box

```{r}
gc_crop <- crop(gc_reproj, bounds)
```

### Resample using the aggregate() function
aggregate exists in baser, and others. Definitely specify that you want the raster:: version.

```{r}
gc_agg <- raster::aggregate(gc_crop, fact = 10)

plot(gc_agg)
```

### Now, let's get safe and sound in ggplot:

First, convert data to a data frame:

```{r}
# xy = df makes lat long a separate column in the data frame!
gc_df <- as.data.frame(gc_agg, xy = TRUE)

# ggplot does not, by default, consider projection
# Use coord_quickmap() to quickly augment this


ggplot(data = gc_df, aes(x = x, y = y)) +
  geom_raster(aes(fill = gc_dem)) +
  coord_quickmap() +
  theme_minimal() +
  scale_fill_gradientn(colors = c(
    "turquoise",
    "magenta",
    "firebrick",
    "white",
    "black",
    "cyan",
    "lightgreen",
    "darkred"
  ))
```

### How can I just select cells that match a given criteria?

```{r}
# First, make a copy of the cropped raster data

gc_hab <- gc_crop

# So, to keep values within a range (1000 - 1500), we gotta make cells outside of that range as NA

# Vertical line = OR

# From gc_hab, anything thats >1500 OR <1000, assign those to NA
gc_hab[gc_hab > 1500 | gc_hab < 1000] <- NA

plot(gc_hab)

```


### Now, let's make this interactive with tmap:

```{r}
tmap_mode("view")

tm_shape(gc_hab) + 
  tm_raster(legend.show = FALSE,
            palette = "plasma")
```

### Kriging rain in Kansas!

Read in KS counties shapefile data:


```{r}
ks_counties <- read_sf(here("data", "ks_counties", "ks_counties_shapefile.shp")) %>% 
  clean_names()

plot(ks_counties)

# Plot can tell you if things are read in properly

# Now, check the CRS within sf:

st_crs(ks_counties)

# Hey! There is none. Let's set one...

st_crs(ks_counties) <- 4326

st_crs(ks_counties)

# WOOOoooooo.....

plot(ks_counties)
```

Practice in ggplot()

```{r}
ggplot(data = ks_counties) +
  geom_sf() +
  theme_minimal()
```

Now, let's read in the rainfall data:

```{r}
ks_rain <- read_csv(here("data", "ks_rain.csv")) %>% 
  clean_names()


```

r doesn't know that this is spatial data yet. Let's get it to recognize this as spatial points. Let's convert it


```{r}
ks_sf <- st_as_sf(ks_rain, coords = c("lon", "lat"), crs = 4326)

# You can set the crs right in the 

plot(ks_sf)
```
let's plot this on top of the counties:


```{r}
ggplot() +
  geom_sf(data = ks_counties, 
          aes(fill = "lightred"),
          show.legend = FALSE) +
  geom_sf(data = ks_sf,
          aes(color = amt,
              size = amt),
          show.legend = FALSE) +
  theme_bw()
```

## Kriging to predict rainfall:

careful here...we're using functions to different packages. gstat doesn't like sf...so we needa convert a whole bunch of stuff.

```{r}
ks_sp <- as_Spatial(ks_sf)

```


### Make a spatial pixels grid that we'll make predictions over...

```{r}
# This gives you min endpoints for a grid
bbox(ks_sp)

# Then make a vector with a bunch of points within that grid...
lat <- seq(37, 40, length.out = 200)
lon <- seq(-94.6, -102, length.out = 200)

# Then, turn this into a spatial grid

grid <- expand.grid(lon = lon, lat = lat)

# This is a 200x200 list of pixels. Woo! Now, lets convert it to spatial...

grid_sf <- st_as_sf(grid, coords = c("lon", "lat"), crs = 4326)

# But, kriging functions don't recognize sf class. so...

grid_sp <- as_Spatial(grid_sf)

plot(grid_sf)

```


### Make a variogram

```{r}
ks_vgm <- variogram(amt ~ 1, data = ks_sp)

plot(ks_vgm)

# Guess your nugget! ~ 0.1
# Guess your sill! ~ 1
# Guess your range! ~ 200

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.1, psill = 1, range = 200, model = "Sph"))

# Fit different models! Spherical = Sph, exponential = Exp, gaussian = Gau

plot(ks_vgm, ks_vgm_fit)

ks_vgm_fit

```

### Now, Krige!

```{r}
ks_krige <- krige(amt ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)

view(ks_krige@data)

# Kriging gives you a predicted amount, plus the variance associated with that prediction

spplot(ks_krige, "var1.pred")
```

# Make a df of krige predictions:

```{r}
ks_df <- data.frame(ks_krige@data["var1.pred"], 
                    ks_krige@data["var1.var"], 
                    ks_krige@coords) %>% 
  rename(longitude = coords.x1, 
         latitude = coords.x2)

# Let's rename these columns...

# Then convert this to an sf object

rain_sf <- st_as_sf(ks_df, coords = c("longitude", "latitude"), crs = 4326)

plot(rain_sf)
```

```{r}
ggplot(rain_sf) +
  geom_sf(aes(color = var1.pred))

```

This is no longer kansas! Let's crop it...

```{r}
# First read in a geom of kansas...
ks <- read_sf(here("data", "states"), layer = "cb_2017_us_state_20m") %>% 
  dplyr::select(NAME) %>% 
  filter(NAME == "Kansas") %>% 
  st_transform(crs = 4326)

plot(ks)

```

```{r}
# Find the intersection of the two...

rain_sf_ks <- st_intersection(rain_sf, ks)

plot(rain_sf_ks)


ggplot(rain_sf_ks) +
  geom_sf(aes(color = var1.pred)) +
  scale_color_gradientn(colors = c("white","yellow","magenta","purple")) +
  theme_minimal()
```

